SUPPORTING INFORMATION \\ Determining sequence-dependent DNA oligonucleotide hybridization and dehybridization mechanisms using coarse-grained molecular simulation, Markov state models, and infrared spectroscopy

### 2021-05-19 SI.tex
Andrei edits
Andy pre-submission edits

%include hyperlinks to figures
%\documentclass{article}
\usepackage{graphicx}
\usepackage[%  
    colorlinks=false,
    pdfborder={0 0 0},
    linkcolor=blue
]{hyperref}

@@

\newcommand*{\rood}[1]{{\color{red}{#1}}}
%\newcommand*{\rood}[1]{#1}
\newcommand*{\blauw}[1]{\textcolor{blue}{#1}}
%\newcommand*{\blauw}[1]{#1}
\newcommand*{\groen}[1]{\textcolor{green}{#1}}
%\newcommand*{\groen}[1]{#1}
%\newcommand*{\blauwr}[1]{{\color{blue}{#1}}}
%\newcommand*{\blauwr}[1]{#1}

\newcommand*{\addref}[1]{\textcolor{red}{\{ADD REF: #1\}}}

\newcommand*{\noter}[1]{\textcolor{red}{[[#1]]}}		% notes on
%\newcommand*{\noter}[1]{}					% notes off

%\usepackage[draft]{todonotes}   % notes shown
\usepackage[disable]{todonotes} % notes hidden

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Meta-data block
%% ---------------
%% Each author should be given as a separate \author command.
%%
%% Corresponding authors should have an e-mail given after the author
%% name as an \email command. Phone and fax numbers can be given
%% using \phone and \fax, respectively; this information is optional.
%%
%% The affiliation of authors is given after the authors; each
%% \affiliation command applies to all preceding authors not already
%% assigned an affiliation.
%%
%% The affiliation takes an option argument for the short name.  This
%% will typically be something like "University of Somewhere".
%%
%% The \altaffiliation macro should be used for new address, etc.
%% On the other hand, \alsoaffiliation is used on a per author basis
%% when authors are associated with multiple institutions.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\author{Michael S. Jones}
\affiliation{%
  Pritzker School of Molecular Engineering, %
  The University of Chicago, %
  929 East 57th Street, Chicago, Illinois 60637, United States%
}

@@

% Fig. S2
\begin{figure}[ht!]
	\centering
    \includegraphics[width=0.8\textwidth]{FigS2.pdf}
    \caption{Convergence of the MSM implied time scales $t_i$ as a function of lag time $\tau$. Solid lines indicate maximum likelihood result while dashed lines show the Bayesian ensemble means. The implied time scales for all sequences converge at a lag time of $\tau$ = 1.2 ns (vertical line). The black solid curve marks equality of the implied time scale and lag time and delimits the shaded region wherein the implied time scales are shorter than the lag time and cannot be resolved.}
    \label{fig:SIFig2}
\end{figure}

@@

% Fig. S3
\begin{figure}[ht!]
	\begin{center}
        \includegraphics[width=1.0\textwidth]{FigS3.pdf}
        \caption{Chapman-Kolmogorov (CK) tests comparing the probabilities of remaining within each macrostates for each sequence as a function of lag time predicted by $k$ applications of an MSM constructed at the $\tau$ = 1.2 ns lag time $\mathbf{P}^k(\tau)$ (dashed blue line) versus those computed from an MSM constructed at that particular lag time $\mathbf{P}(k \tau)$ (solid black line). The good agreement between these two results provides numerical validation of the Markovian nature of the $\tau$ = 1.2 ns lag time MSM.}
        \label{fig:SIFig3}
	\end{center}
\end{figure}

@@

\begin{figure}[ht!]
	\centering
    \caption{GC-mix hybridizes by nucleation-zippering and dehybridizes by fraying-peeling. Tracking of the 10 intermolecular distances between native WC base pairs over the course of four additional hybridization and dissociation events. Symmetrically permutable distances (e.g., A$_1$:T$_{10}$ and T$_{10}$:A$_1$) are reflected across the x-axis to avoid congestion in the plot. Circles superposed on the x-axis indicate the instantaneous MSM state assignment as dissociated D (blue) or hybridized H (orange). Hybridization tends to occur by a nucleation-zippering mechanism, wherein a native G:C WC pair and adjacent A:T pair or central A:T pairs first form prior to rapid formation of the complete duplex. Dehybridization tends to occur by a fraying-peeling mechanism wherein fraying of the two-base AT-tails on one or both sides of the duplex precedes dissociation of the central native base pairs and complete dissolution of the duplex.}
    \label{fig:SIFig5}
\end{figure}

@@

DNA duplex hybridization thermodynamics are most commonly predicted using nearest-neighbor (NN) models where the stability of a given base pair is assumed to solely depend on the identity and orientation of its adjacent base pairs. We apply the NN model reported by SantaLucia\citep{SantaLucia1998AThermodynamics}  to estimate the stability of fully intact, shifted, and frayed duplex configurations observed in our Markov State Model.

The NN model assumes hybridization proceeds in a two-state all-or-nothing manner. The Gibbs free energy difference between the hybridized and dissociated states can be determined from calculation of NN enthalpy and entropy contributions,
\begin{equation}\label{SIe1}
\Delta G_{NN}^{\circ}(T) = \Delta H_{NN}^{\circ} - T\Delta S_{NN}^{\circ}.
\end{equation}
$\Delta H_{NN}^{\circ}$ and $\Delta S_{NN}^{\circ}$ are assumed to be independent of temperature $T$, and are computed from the sum over all NN terms in a given configuration,
\begin{equation}\label{SIe2}
\Delta H_{NN}^{\circ} = \sum_{i}^{n_{NN}}\Delta H_{NN, i}^{\circ} + \sum_{j}^{n_{DE}}\Delta H_{DE, j}^{\circ} + \Delta H_{init}^{\circ},
\end{equation}
\begin{equation}\label{SIe3}
\Delta S_{NN}^{\circ} = \sum_{i}^{n_{NN}}\Delta S_{NN, i}^{\circ} + \sum_{j}^{n_{DE}}\Delta S_{DE, j}^{\circ} + \Delta S_{init}^{\circ} + \Delta S_{sym}^{\circ}.
\end{equation}                          
$\Delta H_{NN, i}^{\circ}$ is the NN enthalpy for a given dinucleotide step within duplex DNA, $\Delta H_{DE, j}^{\circ}$ corresponds to the enthalpic contribution from a given dangling end (DE) base next to the duplex\citep{Santalucia2004TM}, and $\Delta H_{init}^{\circ}$ is a duplex initiation term that accounts for terminal effects on duplex enthalpy. A schematic illustration of the NN and DE contributions for a fully hybridized duplex and an out-of-register shifted state is presented in Fig.~\ref{fig:SIFig6}. $\Delta S_{NN}^{\circ}$ is constructed in an analogous manner from NN, dangling end, and initiation terms, plus a symmetry term, $\Delta S_{sym}^{\circ}$, that applies an entropic penalty to self-complementary sequences for maintaining C2 symmetry.

@@

\bibliography{references_210505}

>>>>

%include hyperlinks to figures
%\documentclass{article}
\usepackage{graphicx}
\usepackage[%  
    colorlinks=true,
    pdfborder={0 0 0},
    linkcolor=black,
    citecolor=black,
    urlcolor=black
]{hyperref}

@@

%\newcommand*{\rood}[1]{{\color{red}{#1}}}
%\newcommand*{\rood}[1]{#1}
%\newcommand*{\blauw}[1]{\textcolor{blue}{#1}}
\newcommand*{\blauw}[1]{#1}
%\newcommand*{\groen}[1]{\textcolor{green}{#1}}
%\newcommand*{\groen}[1]{#1}
%\newcommand*{\blauwr}[1]{{\color{blue}{#1}}}
%\newcommand*{\blauwr}[1]{#1}

%\newcommand*{\addref}[1]{\textcolor{red}{\{ADD REF: #1\}}}

%\newcommand*{\noter}[1]{\textcolor{red}{[[#1]]}}		% notes on
%\newcommand*{\noter}[1]{}					% notes off

%\usepackage[draft]{todonotes}   % notes shown
%\usepackage[disable]{todonotes} % notes hidden

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Meta-data block
%% ---------------
%% Each author should be given as a separate \author command.
%%
%% Corresponding authors should have an e-mail given after the author
%% name as an \email command. Phone and fax numbers can be given
%% using \phone and \fax, respectively; this information is optional.
%%
%% The affiliation of authors is given after the authors; each
%% \affiliation command applies to all preceding authors not already
%% assigned an affiliation.
%%
%% The affiliation takes an option argument for the short name.  This
%% will typically be something like "University of Somewhere".
%%
%% The \altaffiliation macro should be used for new address, etc.
%% On the other hand, \alsoaffiliation is used on a per author basis
%% when authors are associated with multiple institutions.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\author{Michael S. Jones}
\affiliation{%
  Pritzker School of Molecular Engineering, %
  The University of Chicago, %
  5640 South Ellis Avenue, Chicago, Illinois 60637, United States%
}

@@

% Fig. S2
\begin{figure}[ht!]
	\centering
    \includegraphics[width=0.8\textwidth]{FigS2.pdf}
    \caption{Convergence of the MSM implied time scales $t_i$ as a function of lag time $\tau$. Solid lines indicate maximum likelihood result while dashed lines show the Bayesian ensemble means superposed on shaded areas representing 95\% confidence intervals. The high simulation data volume used in our MSM construction means that the maximum likelihood and Bayesian ensemble means are in good agreement and confidence intervals are narrow. The implied time scales for all sequences converge at a lag time of $\tau$ = 1.2 ns (vertical line). The black solid curve marks equality of the implied time scale and lag time and delimits the shaded region wherein the implied time scales are shorter than the lag time and cannot be resolved.}
    \label{fig:SIFig2}
\end{figure}

@@

% Fig. S3
\begin{figure}[ht!]
	\begin{center}
        \includegraphics[width=1.0\textwidth]{FigS3.pdf}
        \caption{Chapman-Kolmogorov (CK) tests comparing the probabilities of remaining within each macrostates for each sequence as a function of lag time predicted by $k$ applications of an MSM constructed at the $\tau$ = 1.2 ns lag time $\mathbf{P}^k(\tau)$ (dashed blue line) versus those computed from an MSM constructed at that particular lag time $\mathbf{P}(k \tau)$ (solid black line). The blue shaded area indicates the estimated standard deviation in $\mathbf{P}^k(\tau)$. The good agreement between these two results provides numerical validation of the Markovian nature of the $\tau$ = 1.2 ns lag time MSM.}
        \label{fig:SIFig3}
	\end{center}
\end{figure}

@@

\begin{figure}[ht!]
	\centering
    \caption{GC-mix hybridizes by nucleation-zippering and dehybridizes by fraying-peeling. Tracking of the 10 intermolecular distances between native WC base pairs over the course of four additional (a) hybridization and (b) dehybridization events. Symmetrically permutable distances (e.g., A$_1$:T$_{10}$ and T$_{10}$:A$_1$) are reflected across the x-axis to avoid congestion in the plot. Circles superposed on the x-axis indicate the instantaneous MSM state assignment as dissociated D (blue) or hybridized H (orange). Hybridization tends to occur by a nucleation-zippering mechanism, wherein a native G:C pair and adjacent A:T pair or 2-3 central A:T pairs first form prior to rapid formation of the duplex. Dehybridization tends to occur by a fraying-peeling mechanism wherein fraying of the two-base AT-tails on one or both sides of the duplex precedes dissociation of the central native base pairs and complete dissolution of the duplex.}
    \label{fig:SIFig5}
\end{figure}

@@

DNA duplex hybridization thermodynamics are commonly predicted using nearest-neighbor (NN) models wherein the stability of a given base pair is assumed to solely depend on the identity and orientation of its adjacent base pairs. We apply the popular NN model reported by SantaLucia\citep{SantaLucia1998AThermodynamics} to estimate the stability of fully intact, shifted, and frayed duplex configurations observed in our MSMs.

The NN model assumes hybridization proceeds in a two-state all-or-nothing manner. The Gibbs free energy difference between the hybridized and dissociated states can be determined from calculation of NN enthalpy and entropy contributions,
\begin{equation}\label{SIe1}
\Delta G_{NN}^{\circ}(T) = \Delta H_{NN}^{\circ} - T\Delta S_{NN}^{\circ}.
\end{equation}
$\Delta H_{NN}^{\circ}$ and $\Delta S_{NN}^{\circ}$ are assumed to be independent of temperature $T$, and are computed from the sum over all NN terms in a given configuration,
\begin{equation}\label{SIe2}
\Delta H_{NN}^{\circ} = \sum_{i}^{n_{NN}}\Delta H_{NN, i}^{\circ} + \sum_{j}^{n_{DE}}\Delta H_{DE, j}^{\circ} + \Delta H_{init}^{\circ},
\end{equation}
\begin{equation}\label{SIe3}
\Delta S_{NN}^{\circ} = \sum_{i}^{n_{NN}}\Delta S_{NN, i}^{\circ} + \sum_{j}^{n_{DE}}\Delta S_{DE, j}^{\circ} + \Delta S_{init}^{\circ} + \Delta S_{sym}^{\circ}.
\end{equation}                          
$\Delta H_{NN, i}^{\circ}$ is the NN enthalpy for a given dinucleotide step within duplex DNA, $\Delta H_{DE, j}^{\circ}$ corresponds to the enthalpic contribution from a given dangling end (DE) base next to the duplex\citep{Santalucia2004TheMotifs}, and $\Delta H_{init}^{\circ}$ is a duplex initiation term that accounts for terminal effects on duplex enthalpy. A schematic illustration of the NN and DE contributions for a fully hybridized duplex and an out-of-register shifted state is presented in Fig.~\ref{fig:SIFig6}. $\Delta S_{NN}^{\circ}$ is constructed in an analogous manner from NN, dangling end, and initiation terms, plus a symmetry term, $\Delta S_{sym}^{\circ}$, that applies an entropic penalty to self-complementary sequences for maintaining C2 symmetry.

@@

\bibliography{references_210514_abv}
